---
title: "Group_9_Analysis"
author: "Brent Strong, Enyu Li, Haotian Wang, Honjin Ren, Mu He"
date: "3/7/2022"
output: pdf_document
---
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, comment = NA, message = FALSE, warning = FALSE)
```

```{r libraries, echo = FALSE}

#Load necessary libraries for the data exploration and analysis

library(tidyverse)
library(moderndive)
library(skimr)
library(kableExtra)
library(gridExtra)
library(broom)
library(olsrr)
library(gapminder)
library(sjPlot)
library(stats)
library(jtools)
library(MASS)
library(janitor)
library(ggplot2)
library(caret)
```

```{r import, echo = FALSE}

#Read in data from github and abbreviate certain country names.

analysis <- read_csv("https://raw.githubusercontent.com/brent-strong/DAS2022-Group-09/main/dataset9.csv")

glimpse(analysis)

for(i in 1:nrow(analysis)){
  if(str_detect(analysis$country_of_origin[i], "Puerto Rico")){
    analysis$country_of_origin[i] <- "Puerto Rico"
  }
  if(str_detect(analysis$country_of_origin[i], "Hawaii")){
    analysis$country_of_origin[i] <- "Hawaii"
  }
  if(str_detect(analysis$country_of_origin[i], "Tanzania")){
    analysis$country_of_origin[i] <- "Tanzania"
  }
  else{
    analysis$country_of_origin[i] <- analysis$country_of_origin[i]
  }
}

# View(analysis)

```

```{r , echo = FALSE}
# glimpse(analysis)
#Examine the proportion of batches that are good:
mean(analysis$Qualityclass=="Good")
```

```{r, echo=FALSE, eval = TRUE}
my_skim2 <- skim_with(numeric = sfl(hist = NULL))
analysis %>%
  dplyr::select(-country_of_origin, -Qualityclass) %>%
  my_skim2() %>%
  dplyr::select(-c(n_missing, complete_rate, skim_type)) %>%
  kable(col.names = c("Variable", "Mean", "SD", "Min.", "1st Q.", "Median",
                        "3rd Q.", "Max."), 
        caption = 'Summary statistics of continuous variables.',
        booktabs = TRUE, format = "latex", digits = 2) %>%
  kable_styling(font_size = 9, latex_options = "HOLD_position")
```

```{r, eval = TRUE}
my_skim <- skim_with(base = sfl(n = length))
my.analysis <- analysis %>%
  mutate(Qualityclassindicator = as.numeric(Qualityclass=="Good"))
my.analysis %>%
  group_by(country_of_origin) %>%
  dplyr::select(country_of_origin, Qualityclassindicator) %>%
  my_skim() %>%
  dplyr::select(country_of_origin, n, numeric.mean) %>%
  transmute(country_of_origin=country_of_origin,
            number_of_batch=n,
            Proportion_of_good_quality=numeric.mean) %>%
  kable(caption = 'Number of batch and proportion of good quality by country of origin', booktabs = TRUE, linesep = "", digits = 2) %>%
  kable_styling(font_size = 10, latex_options = "HOLD_position")
```

```{r, eval = TRUE}
analysis %>%
  group_by(Qualityclass) %>%
  dplyr::select(Qualityclass, aroma, flavor, acidity, category_two_defects) %>%
  my_skim() %>%
  transmute(Variable=skim_variable, Qualityclass=Qualityclass, n=n, Mean=numeric.mean, SD=numeric.sd,
            Min=numeric.p0, Median=numeric.p50,  Max=numeric.p100,
            IQR = numeric.p75-numeric.p50) %>%
  kable(caption = 'Summary statistics of continuous variables by quality of coffee', booktabs = TRUE, linesep = "", digits = 2) %>%
  kable_styling(font_size = 10, latex_options = "HOLD_position")
```

```{r aroma, echo=FALSE, fig.width = 13, fig.align = "center", fig.cap = "Histogram and boxplot for aroma.", fig.pos = 'H', message = FALSE}
#Generate histogram to visualize aroma data.
p1 <- ggplot(data = analysis, mapping = aes(x = aroma)) +
  geom_histogram()
#Generate boxplot to visualize aroma data.
p2 <- ggplot(data = analysis, mapping = aes(y = aroma)) +
  geom_boxplot()
grid.arrange(p1, p2, ncol = 2)
#Single observation with a value of 0 is odd.
```

```{r flavor, echo=FALSE, fig.width = 13, fig.align = "center", fig.cap = "Histogram and boxplot for flavor.", fig.pos = 'H', message = FALSE}
#Generate histogram to visualize flavor data.
p1 <- ggplot(data = analysis, mapping = aes(x = flavor)) +
  geom_histogram()
#Generate boxplot to visualize flavor data.
p2 <- ggplot(data = analysis, mapping = aes(y = flavor)) +
  geom_boxplot()
grid.arrange(p1, p2, ncol = 2)
# Single observation with a value of 0 is odd.
```


```{r acidity, echo=FALSE, fig.width = 13, fig.align = "center", fig.cap = "Histogram and boxplot for acidity.", fig.pos = 'H', message = FALSE}
#Generate histogram to visualize acidity data.
p1 <- ggplot(data = analysis, mapping = aes(x = acidity)) +
  geom_histogram()
#Generate boxplot to visualize acidity data.
p2 <- ggplot(data = analysis, mapping = aes(y = acidity)) +
  geom_boxplot()
grid.arrange(p1, p2, ncol = 2)

# Single observation with a value of 0 is odd. See if this same observation has a value of 0 for flavor and aroma as well.
analysis %>% 
  filter(acidity == 0 & flavor == 0 & aroma == 0)
#Will delete the observation in the formal analysis. 

```

```{r category_two_defects, echo=FALSE, fig.width = 13, fig.align = "center", fig.cap = "Histogram and boxplot for category two defects.", fig.pos = 'H', message = FALSE}
#Generate histogram to visualize category_two_defects data.
p1 <- ggplot(data = analysis, mapping = aes(x = category_two_defects)) +
  geom_histogram()
#Generate boxplot to visualize category_two_defects data.
p2 <- ggplot(data = analysis, mapping = aes(y = category_two_defects)) +
  geom_boxplot()
grid.arrange(p1, p2, ncol = 2)
```

Due to the right-skewed distribution of category_two_defects, log-transformation is applied.

```{r category_two_defects2, echo = FALSE}

# Apply log(x+1) transformation and then visualize the data set.

analysis1 <- analysis %>%
  mutate(defects_log = log(category_two_defects+1))
```

``` {r category_two_defects3, echo=FALSE, fig.width = 13, fig.align = "center", fig.cap = "Histogram and boxplot for category two defects after log transformation.", fig.pos = 'H', message = FALSE}
# Generate histogram to visualize category_two_defects data after log-transformation.
p1 <- ggplot(data = analysis1, mapping = aes(x = defects_log)) +
  geom_histogram()
# Generate boxplot to visualize category_two_defects data after log-transformation.
p2 <- ggplot(data = analysis1, mapping = aes(y = defects_log)) +
  geom_boxplot()
grid.arrange(p1, p2, ncol = 2)
```

``` {r , echo=FALSE, fig.width = 13, fig.align = "center", fig.cap = "Histogram and boxplot for altitude.", fig.pos = 'H', message = FALSE}
# Generate histogram to visualize category_two_defects data after log-transformation.
p1 <- ggplot(data = analysis, mapping = aes(x = altitude_mean_meters)) +
  geom_histogram()
# Generate boxplot to visualize category_two_defects data after log-transformation.
p2 <- ggplot(data = analysis, mapping = aes(y = altitude_mean_meters)) +
  geom_boxplot()
grid.arrange(p1, p2, ncol = 2)
```
Mt. Everest is only 8,849 meters tall. Remove any observations with altitudes higher than that.

```{r altitude1, echo = FALSE}

# Mt. Everest is only 8,849 meters tall. Remove any observations with altitudes higher than that.

analysis2 <- analysis %>%
  filter(altitude_mean_meters < 8849)
```

``` {r altitude2, echo=FALSE, fig.width = 13, fig.align = "center", fig.cap = "Histogram and boxplot for altitude after removing implausuble observations.", fig.pos = 'H', message = FALSE}
# Generate histogram to visualize category_two_defects data after log-transformation.
p1 <- ggplot(data = analysis2, mapping = aes(x = altitude_mean_meters)) +
  geom_histogram()
# Generate boxplot to visualize category_two_defects data after log-transformation.
p2 <- ggplot(data = analysis2, mapping = aes(y = altitude_mean_meters)) +
  geom_boxplot()
grid.arrange(p1, p2, ncol = 2)
```

``` {r harvested, echo=FALSE, fig.width = 13, fig.align = "center", fig.cap = "Histogram and boxplot for harvested.", fig.pos = 'H', message = FALSE}
# Generate histogram to visualize category_two_defects data after log-transformation.
p1 <- ggplot(data = analysis, mapping = aes(x = harvested)) +
  geom_histogram()
# Generate boxplot to visualize category_two_defects data after log-transformation.
p2 <- ggplot(data = analysis, mapping = aes(y = harvested)) +
  geom_boxplot()
grid.arrange(p1, p2, ncol = 2)
```

```{r country, echo = FALSE}

# glimpse(analysis)

# Create a new data set with the proportion of batches that were good and the number of observations.  

analysis3 <- analysis %>%
  mutate(Qualityclass_indicator = 0 + as.numeric(Qualityclass=="Good")) %>%
  group_by(country_of_origin) %>%
  summarize(proportion = mean(Qualityclass_indicator), n=n()) %>%
  arrange(desc(proportion))

# View(analysis3)

#Create a bar chart to visualize the proportion of batches that were good for each country. Split up the visualizations into four groups. 

p1<-ggplot(data = analysis3[1:9,], mapping = aes(x = country_of_origin, y = proportion)) +
  geom_col() + geom_text(data = analysis3[1:9,], aes(label = n), nudge_y=0.25)

p2<-ggplot(data = analysis3[10:18,], mapping = aes(x = country_of_origin, y = proportion)) +
  geom_col() + geom_text(data = analysis3[10:18,], aes(label = n), nudge_y=0.35)

p3<-ggplot(data = analysis3[19:26,], mapping = aes(x = country_of_origin, y = proportion)) +
  geom_col() + geom_text(data = analysis3[19:26,], aes(label = n), nudge_y=0.35)

p4<-ggplot(data = analysis3[27:34,], mapping = aes(x = country_of_origin, y = proportion)) +
  geom_col() + geom_text(data = analysis3[27:34,], aes(label = n), nudge_y=0.3)

grid.arrange(p1, p2, p3, p4, nrow = 4)
```

```{r boxplots, echo = FALSE}

# glimpse(analysis)

# Create the final data set. Remove observations that have implausible values for at least one of the variables. Standardize aroma, flavor, acidity, and altitude. Apply a log(x+1) transformation to category_two_defects variable. Turn Qualityclass into an indicator variable so that it can be used in the logistic regression analysis. Remove unneeded variables.

coffee_final <- analysis %>%
  filter(acidity != 0 & altitude_mean_meters < 8849 & !is.na(harvested)) %>%
  mutate(aroma = as.numeric(scale(aroma)), flavor=as.numeric(scale(flavor)), acidity=as.numeric(scale(acidity)), defects_log = log(category_two_defects + 1), altitude = as.numeric(scale(altitude_mean_meters)), year = as.factor(harvested), Qualityclass = 0 + as.numeric(Qualityclass == "Good")) 

##divide 3 levels for the altitude @hameed_coffee_2020
coffee_final$level_1 <- ifelse(coffee_final$altitude_mean_meters<900,1,0)
coffee_final$level_2 <- ifelse(coffee_final$altitude_mean_meters>=900 & coffee_final$altitude_mean_meters<=1200,1,0)
coffee_final$level_3 <- ifelse(coffee_final$altitude_mean_meters>1200,1,0)
coffee_final <- coffee_final %>%
  mutate(level = as.character(level_1 + level_2*2 + level_3*3))

# Finally we can get the final dataset coffee_final

coffee_final <- coffee_final %>%
  dplyr::select(country_of_origin, aroma, flavor, acidity, defects_log, year, level, Qualityclass)

# View the final data set and make that the standardizations and transformation were appropriately applied.

glimpse(coffee_final)
# View(coffee_final)
#skim_without_charts(coffee_final)

#Examine the number of observations by year.

coffee_final %>%
  group_by(year) %>%
  summarize(n=n())

```

```{r}

# We generate a table to have a look at the proportion and counts for both Good and Poor quality based on the country of origin and harvested years respectively.

coffee_final %>% 
  tabyl(Qualityclass, country_of_origin) %>% 
  adorn_percentages() %>% 
  adorn_pct_formatting() %>% 
  adorn_ns()

coffee_final %>% 
  tabyl(Qualityclass, year) %>% 
  adorn_percentages() %>% 
  adorn_pct_formatting() %>% 
  adorn_ns()

```

```{r}

# Plot a bar chart to get intuition of whether the quality of coffee is influenced by the year of harvesting.

ggplot(coffee_final, aes(x= year,  y = ..prop.., group=Qualityclass, fill=Qualityclass)) + 
    geom_bar(position="dodge", stat="count") +
    labs(y = "Proportion")

```


# Formal Analysis

## Build the models


**Only base on the altitude**
```{r}
model_level <- glm(Qualityclass ~ level - 1, data = coffee_final,family = binomial(link = "logit"))
summary(model_level)
```

**Base on the year of harvest**
```{r}
model_year <- glm(Qualityclass ~ year, data = coffee_final,family = binomial(link = "logit"))
summary(model_year)
```

**Base on the country**
```{r}
model_country <- glm(Qualityclass ~ country_of_origin, data = coffee_final,family = binomial(link = "logit"))
summary(model_country)
```

According the result before, we choose some significant country as a class variable.
```{r}
coffee_final$Colombia <- ifelse(coffee_final$country_of_origin == 'Colombia',1,0)
coffee_final$Mexico <- ifelse(coffee_final$country_of_origin == 'Mexico',1,0)
coffee_final$Honduras <- ifelse(coffee_final$country_of_origin == 'Honduras',1,0)
coffee_final$Kenya <- ifelse(coffee_final$country_of_origin == 'Kenya',1,0)
```

```{r}
model_co_4 <- glm(Qualityclass ~ Colombia + Mexico + Honduras + Kenya-1, data = coffee_final,family = binomial(link = "logit"))
summary(model_co_4)
```

**Base on the year and country**
```{r}
model_cn_ye <- glm(Qualityclass ~ country_of_origin + year, data = coffee_final,family = binomial(link = "logit"))
summary(model_cn_ye)
```

**Base on the altitude and country**
```{r,warning=FALSE}
model_al_co <- glm(Qualityclass ~ level + Colombia + Mexico + Honduras + Kenya, data = coffee_final,family = binomial(link = "logit"))
summary(model_al_co)
```


**Base on the 3**
```{r,warning=FALSE}
model_al_co <- glm(Qualityclass ~ level + country_of_origin + year - 1, data = coffee_final,family = binomial(link = "logit"))
summary(model_al_co)
```


Colombia + Mexico + Honduras + Kenya


**Consider everything**
```{r}
model_all <- glm(Qualityclass ~ aroma + flavor + acidity + country_of_origin + defects_log + level + year, data = coffee_final,family = binomial(link = "logit"))
summary(model_all)
```

```{r}
model_all_2 <- glm(Qualityclass ~ aroma + flavor + acidity + Colombia + Mexico + Honduras + Kenya + defects_log + level + year, data = coffee_final, family = binomial(link = "logit"))
summary(model_all_2)
```

```{r}
stepAIC(model_all_2, direction = 'both')
```

```{r}
model_best <- glm(Qualityclass ~ aroma + flavor + acidity + Colombia + Mexico + defects_log, data = coffee_final,family = binomial(link = "logit"))
summary(model_best)
```

```{r}
set.seed(9)
folds <- createFolds(y=coffee_final$Qualityclass, k=10)
accuracy <- as.numeric()
sensitivity <- as.numeric()
specificity <- as.numeric()
for(i in 1:10){
  fold_test <- coffee_final[folds[[i]],]
  fold_train <- coffee_final[-folds[[i]],]
  fold_pre <- glm(Qualityclass ~ aroma + flavor + acidity + Colombia + Mexico + defects_log,family = binomial(link = logit), data =fold_train )
  fold_predict <- predict(fold_pre,type='response',newdata=fold_test)
  fold_predict <- ifelse(fold_predict >= 0.5, 1, 0)
  accuracy[i] <- mean(fold_predict == fold_test[,8])
  sensitivity[i] <- sum(fold_predict + fold_test[,8] == 2) / sum(fold_test[,8] == 1)
  specificity[i] <- sum(fold_predict + fold_test[,8] == 0) / sum(fold_test[,8] == 0)
}
mean(accuracy)

mean(sensitivity)

mean(specificity)
```








