---
title: "G9_p2"
author: "Hongjin Ren"
date: "2022/3/10"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r libraries}
library(tidyverse)
library(moderndive)
library(skimr)
library(kableExtra)
library(gridExtra)
library(broom)
library(olsrr)
library(gapminder)
library(sjPlot)
library(stats)
library(jtools)
```

```{r}
coffee <- read_csv("https://raw.githubusercontent.com/brent-strong/DAS2022-Group-09/main/dataset9.csv")
```

- country_of_origin – Country where the coffee bean originates from.
- aroma – Aroma grade (ranging from 0-10)
- flavor – Flavor grade (ranging from 0-10)
- acidity – Acidity grade (ranging from 0-10)
- category_two_defects – Count of category 2 type defects in the batch of coffee beans tested.
- altitiude_mean_meters – Mean altitude of the growers farm (in metres)
- harvested – Year the batch was harvested
- Qualityclass – Quality score for the batch (Good - ≥ 82.5, Poor - <82.5). Note: 82.5
was selected as the cut off as this is the median score for all the batches tested.

```{r}
co_Na <- na.omit(coffee)
```

```{r}
cor(co_Na[2:6])
```


```{r}
skim(coffee)
cor(coffee[2:6])
```



